---
title: "Hw2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

```{r echo=FALSE, results="hide", fig1, fig.height = 8, fig.width = 7, warning=FALSE,message=FALSE}
library(gridExtra)
library(ggplot2)
library(reshape)
library(dplyr)
library(maps)
library(viridis)
library("rnaturalearth")
library("rnaturalearthdata")
library("sf")

world <- ne_countries(scale = "medium", returnclass = "sf")

###load data
ele <- read.table("./NASA/Files/elevation.dat",header=TRUE)
Lon <- colnames(ele)
Lon <- unlist(lapply(Lon,function(x)gsub("X.","",x)))
names(ele) <- Lon

ele.Lon <- as.numeric(Lon)
ele.Lat <- as.numeric(rownames(ele))

metric.Lon <- scan("./NASA/Files/Lon.txt",what="",sep="\t")
metric.Lon <- unlist(lapply(metric.Lon,function(x) gsub("W","",x)))
metric.Lon <- as.numeric(metric.Lon)

metric.Lat <- scan("./NASA/Files/Lat.txt",what="",sep="\n")
metric.Lat <- unlist(lapply(metric.Lat,function(x){
  x <- gsub("N","",x)
  x <- gsub("S","",x)
})                            )
metric.Lat <- as.numeric(metric.Lat)
metric.Lat[16:24] <- -metric.Lat[16:24]


## average for each month
ozone.month.mean <- data.frame(Lat = rep(metric.Lat,24), 
                               Lon = rep(metric.Lon,each = 24))


## mean of each month
for(yy in 1:12){
  tmp <- NULL
  for(t in seq(yy,72,12)){
    ozone <- read.table(paste("./NASA/Files/ozone",t,".txt",sep=""),header=FALSE,skip=7,na.strings="....")
    tmp <- cbind(tmp,as.vector(as.matrix(ozone[,-(1:3)])))
  }
  ozone.month.mean[[yy+2]] <- apply(tmp,1,mean)
}
names(ozone.month.mean)[3:14] <- c("Jan", "Feb", "Mar", "Apr", "May", "June", 
                                   "July", "Aug", "Sept", "Oct", "Nov", "Dec")

df1 <- melt(ozone.month.mean, id=c("Lat", "Lon"))
names(df1)[4] <- "OzoneAbundance"
df1$Lon <- -df1$Lon

US <- map_data("world")
US <- US[US$region=="USA",]

p1 <- ggplot(data = world) +
  geom_sf()+
  geom_point(dat = df1, aes(x = Lon, y = Lat, alpha = OzoneAbundance, color = OzoneAbundance, size=OzoneAbundance)) +
  ylim(c(-20,37))+
  scale_alpha_continuous(name="OzoneAbundance", range=c(0.1,0.5)) + 
  scale_size_continuous(name="OzoneAbundance", range=c(1,8)) +
  scale_fill_gradient(low = "white",high = "black") +
  facet_wrap(~variable) +
  ggtitle("Ozone month average")+
  theme(axis.text.x = element_text(angle = 90))+
  coord_sf( c(-56, -114))
   # geom_polygon(data = US, aes(x=long, y = lat, group = group), fill="black", alpha=0.5)  +
# scale_color_viridis() +

  
  # geom_point( dat = df1, aes(x=Lon, y=Lat, size=OzoneAbundance,alpha = OzoneAbundance), shape=20, stroke=FALSE) 
  #   scale_size_continuous(name="Population (in M)", trans="log", range=c(1,12), breaks=mybreaks) +
  #   scale_alpha_continuous(name="Population (in M)", trans="log", range=c(0.1, .9), breaks=mybreaks) +
  #   scale_color_viridis(option="magma", trans="log", breaks=mybreaks, name="Population (in M)" )

# p1 <- ggplot(df1, aes(x = Lon, y = Lat)) + 
#   facet_wrap(~variable) + 
#   geom_tile(aes(fill = OzoneAbundance), colour = "white")  + 
#   scale_fill_gradient(low = "white",high = "black") + 
#   ggtitle("Ozone month average")
# 

## average for each year
ozone.year.mean <- data.frame(Lat = rep(metric.Lat,24), 
                               Lon = rep(metric.Lon,each = 24))
## mean of each year
for(yy in 1:6){
  tmp <- NULL
  for(t in ((yy-1)*12+1):((yy-1)*12+12)) {
    ozone <- read.table(paste("./NASA/Files/ozone",t,".txt",sep=""),header=FALSE,skip=7,na.strings="....")
    tmp <- cbind(tmp,as.vector(as.matrix(ozone[,-(1:3)])))
  }
  ozone.year.mean[[yy+2]] <- apply(tmp,1,mean)
}
names(ozone.year.mean)[3:8] <- 1995:2000

df2 <- melt(ozone.year.mean, id=c("Lat", "Lon"))
df2$Lon <- -df2$Lon
names(df2)[4] <- "OzoneAbundance"
# p2 <- ggplot(df2, aes(x = Lon, y = Lat)) + 
#   facet_wrap(~variable) + 
#   geom_tile(aes(fill = OzoneAbundance), colour = "white")  + 
#   scale_fill_gradient(low = "white",high = "black") 

p2 <- ggplot(data = world) +
  geom_sf()+
  geom_point(dat = df2, aes(x = Lon, y = Lat, alpha = OzoneAbundance, color = OzoneAbundance, size=OzoneAbundance)) +
  ylim(c(-20,37))+
  scale_alpha_continuous(name="OzoneAbundance", range=c(0.1,0.5)) + 
  scale_size_continuous(name="OzoneAbundance", range=c(1,8)) +
  scale_fill_gradient(low = "white",high = "black") +
  ggtitle("Yearly trend of the Ozone")+
  facet_wrap(~variable) +
  theme(axis.text.x = element_text(angle = 90))+
  coord_sf( c(-56, -114))


#pdf("q1.pdf", paper="a4")

#dev.off()

```

\newpage
## 2
```{r echo=FALSE, results="hide"}
dat <- read.csv("./Diving2000.csv") # read data
diver_uni <- unique(dat$diver) # all divers
n_diver <- length(diver_uni) #no. of divers
judge_uni <- unique(dat$judge) # all judges
n_judge <- length(judge_uni) #no. of judges
c_diver <- c_judge <- vector()
adj <-  matrix(0, n_diver, n_judge)

for(i in 1:n_diver) {
  tmp <- dat[dat$diver==diver_uni[i],"dcountry"]
  c_diver[i] <- as.character(tmp[1])
}

for(i in 1:n_judge) {
  tmp <- dat[dat$judge==judge_uni[i],"jcountry"]
  c_judge[i] <- as.character(tmp[1])
}

AND1 <- function (...)  Reduce("&", list(...))

for(i in 1:n_diver) {
  for(j in 1:n_judge) {
    idx <- AND1(dat$diver==diver_uni[i], dat$judge == judge_uni[j]) # index
    if(sum(idx)>0) adj[i,j] <- mean(dat[idx,"score"]) # mean score
  }
}

highscore <- matrix(0, n_diver, n_judge)
highscore[adj>quantile(adj,0.9)]<- 1
same_c <-  matrix(0, n_diver, n_judge)
for(i in 1:n_diver) {
  for(j in 1:n_judge) {
    idx <- AND1(dat$diver==diver_uni[i], dat$judge == judge_uni[j]) # index
    tmp1 <- dat[idx,"dcountry"]
    tmp2 <- dat[idx,"jcountry"]
    if(sum(idx)>0) same_c[i,j] <- tmp1[1]==tmp2[1]# mean score
  }
}



heat.df <- as.data.frame(highscore*same_c)
names(heat.df) <- make.unique(c_judge)
row.names(heat.df) <- make.unique(c_diver)
heatmap(as.matrix(heat.df))
#Highscores (score>7.9) + the diver and the judge are from the same country

```

The red part indicates the highscores (score>7.5, the 90\% quantile of the scores) as well as the diver and judge are from the same country to visulize the bias.
It is easy to see that the judges tend to be biased towards divers from their own countries, especially in USA, China, Russia, Canada, and UK. The highest scores of the diver often come from the judges from the same country. 

\newpage
## Appendix
```{r eval=F}
library(gridExtra)
library(ggplot2)
library(reshape)
library(dplyr)
library(maps)
library(viridis)
library("rnaturalearth")
library("rnaturalearthdata")
library("sf")

world <- ne_countries(scale = "medium", returnclass = "sf")

###load data
ele <- read.table("./NASA/Files/elevation.dat",header=TRUE)
Lon <- colnames(ele)
Lon <- unlist(lapply(Lon,function(x)gsub("X.","",x)))
names(ele) <- Lon

ele.Lon <- as.numeric(Lon)
ele.Lat <- as.numeric(rownames(ele))

metric.Lon <- scan("./NASA/Files/Lon.txt",what="",sep="\t")
metric.Lon <- unlist(lapply(metric.Lon,function(x) gsub("W","",x)))
metric.Lon <- as.numeric(metric.Lon)

metric.Lat <- scan("./NASA/Files/Lat.txt",what="",sep="\n")
metric.Lat <- unlist(lapply(metric.Lat,function(x){
  x <- gsub("N","",x)
  x <- gsub("S","",x)
})                            )
metric.Lat <- as.numeric(metric.Lat)
metric.Lat[16:24] <- -metric.Lat[16:24]


## average for each month
ozone.month.mean <- data.frame(Lat = rep(metric.Lat,24), 
                               Lon = rep(metric.Lon,each = 24))


## mean of each month
for(yy in 1:12){
  tmp <- NULL
  for(t in seq(yy,72,12)){
    ozone <- read.table(paste("./NASA/Files/ozone",t,".txt",sep=""),header=FALSE,skip=7,na.strings="....")
    tmp <- cbind(tmp,as.vector(as.matrix(ozone[,-(1:3)])))
  }
  ozone.month.mean[[yy+2]] <- apply(tmp,1,mean)
}
names(ozone.month.mean)[3:14] <- c("Jan", "Feb", "Mar", "Apr", "May", "June", 
                                   "July", "Aug", "Sept", "Oct", "Nov", "Dec")

df1 <- melt(ozone.month.mean, id=c("Lat", "Lon"))
names(df1)[4] <- "OzoneAbundance"
df1$Lon <- -df1$Lon

US <- map_data("world")
US <- US[US$region=="USA",]

p1 <- ggplot(data = world) +
  geom_sf()+
  geom_point(dat = df1, aes(x = Lon, y = Lat, alpha = OzoneAbundance, color = OzoneAbundance, size=OzoneAbundance)) +
  ylim(c(-20,37))+
  scale_alpha_continuous(name="OzoneAbundance", range=c(0.1,0.5)) + 
  scale_size_continuous(name="OzoneAbundance", range=c(1,8)) +
  scale_fill_gradient(low = "white",high = "black") +
  facet_wrap(~variable) +
  ggtitle("Ozone month average")+
  theme(axis.text.x = element_text(angle = 90))+
  coord_sf( c(-56, -114))
   # geom_polygon(data = US, aes(x=long, y = lat, group = group), fill="black", alpha=0.5)  +
# scale_color_viridis() +

  
  # geom_point( dat = df1, aes(x=Lon, y=Lat, size=OzoneAbundance,alpha = OzoneAbundance), shape=20, stroke=FALSE) 
  #   scale_size_continuous(name="Population (in M)", trans="log", range=c(1,12), breaks=mybreaks) +
  #   scale_alpha_continuous(name="Population (in M)", trans="log", range=c(0.1, .9), breaks=mybreaks) +
  #   scale_color_viridis(option="magma", trans="log", breaks=mybreaks, name="Population (in M)" )

# p1 <- ggplot(df1, aes(x = Lon, y = Lat)) + 
#   facet_wrap(~variable) + 
#   geom_tile(aes(fill = OzoneAbundance), colour = "white")  + 
#   scale_fill_gradient(low = "white",high = "black") + 
#   ggtitle("Ozone month average")
# 

## average for each year
ozone.year.mean <- data.frame(Lat = rep(metric.Lat,24), 
                               Lon = rep(metric.Lon,each = 24))
## mean of each year
for(yy in 1:6){
  tmp <- NULL
  for(t in ((yy-1)*12+1):((yy-1)*12+12)) {
    ozone <- read.table(paste("./NASA/Files/ozone",t,".txt",sep=""),header=FALSE,skip=7,na.strings="....")
    tmp <- cbind(tmp,as.vector(as.matrix(ozone[,-(1:3)])))
  }
  ozone.year.mean[[yy+2]] <- apply(tmp,1,mean)
}
names(ozone.year.mean)[3:8] <- 1995:2000

df2 <- melt(ozone.year.mean, id=c("Lat", "Lon"))
df2$Lon <- -df2$Lon
names(df2)[4] <- "OzoneAbundance"
# p2 <- ggplot(df2, aes(x = Lon, y = Lat)) + 
#   facet_wrap(~variable) + 
#   geom_tile(aes(fill = OzoneAbundance), colour = "white")  + 
#   scale_fill_gradient(low = "white",high = "black") 

p2 <- ggplot(data = world) +
  geom_sf()+
  geom_point(dat = df2, aes(x = Lon, y = Lat, alpha = OzoneAbundance, color = OzoneAbundance, size=OzoneAbundance)) +
  ylim(c(-20,37))+
  scale_alpha_continuous(name="OzoneAbundance", range=c(0.1,0.5)) + 
  scale_size_continuous(name="OzoneAbundance", range=c(1,8)) +
  scale_fill_gradient(low = "white",high = "black") +
  ggtitle("Yearly trend of the Ozone")+
  facet_wrap(~variable) +
  theme(axis.text.x = element_text(angle = 90))+
  coord_sf( c(-56, -114))


#pdf("q1.pdf", paper="a4")
grid.arrange(p1, p2, nrow=2)
#dev.off()


dat <- read.csv("./Diving2000.csv") # read data
diver_uni <- unique(dat$diver) # all divers
n_diver <- length(diver_uni) #no. of divers
judge_uni <- unique(dat$judge) # all judges
n_judge <- length(judge_uni) #no. of judges
c_diver <- c_judge <- vector()
adj <-  matrix(0, n_diver, n_judge)

for(i in 1:n_diver) {
  tmp <- dat[dat$diver==diver_uni[i],"dcountry"]
  c_diver[i] <- as.character(tmp[1])
}

for(i in 1:n_judge) {
  tmp <- dat[dat$judge==judge_uni[i],"jcountry"]
  c_judge[i] <- as.character(tmp[1])
}

AND1 <- function (...)  Reduce("&", list(...))

for(i in 1:n_diver) {
  for(j in 1:n_judge) {
    idx <- AND1(dat$diver==diver_uni[i], dat$judge == judge_uni[j]) # index
    if(sum(idx)>0) adj[i,j] <- mean(dat[idx,"score"]) # mean score
  }
}

highscore <- matrix(0, n_diver, n_judge)
highscore[adj>quantile(adj,0.9)]<- 1
same_c <-  matrix(0, n_diver, n_judge)
for(i in 1:n_diver) {
  for(j in 1:n_judge) {
    idx <- AND1(dat$diver==diver_uni[i], dat$judge == judge_uni[j]) # index
    tmp1 <- dat[idx,"dcountry"]
    tmp2 <- dat[idx,"jcountry"]
    if(sum(idx)>0) same_c[i,j] <- tmp1[1]==tmp2[1]# mean score
  }
}



heat.df <- as.data.frame(highscore*same_c)
names(heat.df) <- make.unique(c_judge)
row.names(heat.df) <- make.unique(c_diver)
heatmap(as.matrix(heat.df))
#Highscores (score>7.5) + the diver and the judge are from the same country

```

