---
title: "STAT 3280 HW1"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Q1

The first question

```{r cars, echo=FALSE,fig.align='center', fig.height=10, fig.width=10} 
par(mfrow=c(1,1))
digit = matrix(scan('HandWritten.txt'), 16, byrow = TRUE)
image(t(digit[16:1, ]))
```
Based on the plot, the digit is 3

\newpage
##Q2


```{r pressure, echo=FALSE, fig.align='center', fig.height=10, fig.width=10}
df = read.csv('mixture.csv', row.names = 1)
col_list<-ifelse(df$y==1,'blue','red')
plot(x2 ~ x1, df, cex.lab = 1.5, xlim = c(-3.5, 5.5), ylim = c(-3.5, 5.5), bty = 'L',
     col=col_list)
lines(df$x1, (0.978 + 0.134*df$x1)/1.398, lwd = 2)
x1 = sort(df$x1)
x = c(min(x1)-0.5, x1, max(x1)+0.5)
polygon(c(x[1], x, x[length(x)]),
        c(min(df$x2)-0.5, (0.978 + 0.134*x)/1.398, min(df$x2)-0.5), col = "pink")
polygon(c(x[1], x, x[length(x)]),
        c(max(df$x2)+0.5, (0.978 + 0.134*x)/1.398, max(df$x2)+0.5), col = "lightblue")
pchs = ifelse(df$y == 0, 2, 8)
points(df$x1, df$x2, pch = pchs,col=col_list)
legend('topright', legend = c(0, 1), pch = c(2, 8), cex = 1.5, bty = 'n', inset = 0.01
       ,col=c('red','blue'))
```
\newpage
##Q3

Before the analysis, I divide this data into 2 groups, the old group whose teachers are 40 years or older and the young group whose teachers are less than 40 years old.

```{r warning=FALSE, echo=FALSE, fig.align='center', fig.height=4, fig.width=10}
df = read.csv('TeacherHires.csv', stringsAsFactors = FALSE, na.strings = 'N/A')
df$Calculated.Age = as.numeric(df$Calculated.Age)
df = df[!is.na(df$Calculated.Age), ]
df$Age = ifelse(df$Calculated.Age < 40, 'young', 'old')
df$Interview = ifelse(df$Interview == 'no', 'no', 'yes')
df$Hired = ifelse(df$Hired == 'no', 'no', 'yes')
par(mfrow = c(1, 2), mar = c(3, 3, 2, 3))
tb1 = round(prop.table(table(df$Interview, df$Age), 2), 4)
barplot(tb1, beside = TRUE, legend.text = TRUE, ylim  = c(0, 1), xlab = 'Age',
        ylab = 'Percentage', main = 'Interviewed or not vs Age',
        args.legend = list(bty = 'n'))
text(c(1.5, 2.5), tb1[, 1]+0.05, paste0(tb1[, 1]*100, '%'))
text(c(4.5, 5.5), tb1[, 2]+0.05, paste0(tb1[, 2]*100, '%'))
tb2 = round(prop.table(table(df$Hired, df$Age), 2), 4)
barplot(tb2, beside = TRUE, legend.text = TRUE, ylim  = c(0, 1), xlab = 'Age',
        ylab = 'Percentage', main = 'Hired or not vs Age',
        args.legend = list(bty = 'n'))
text(c(1.5, 2.5), tb2[, 1]+0.05, paste0(tb2[, 1]*100, '%'))
text(c(4.5, 5.5), tb2[, 2]+0.05, paste0(tb2[, 2]*100, '%'))
```

Based on the barplots, there is evidence of age discrimination in the interviewing and hiring of teachers.

```{r warning=FALSE, echo=FALSE, fig.align='center', fig.height=7, fig.width=12}
df$GPA.undergrad = as.numeric(df$GPA.undergrad)
df$GPA.GRAD = as.numeric(df$GPA.GRAD)
df = df[!((is.na(df$GPA.GRAD)) | (is.na(df$GPA.undergrad))), ]
par(mfrow = c(2, 2), mar = c(3, 3, 3, 2))
boxplot(GPA.undergrad ~ Hired, df, main = 'Hired or not vs GPA Undergrad')
boxplot(GPA.GRAD ~ Hired, df, main = 'Hired or not vs GPA Grad')
ind = grep('months', df$Work.Experience )
df$Work.Experience = as.numeric(gsub('\\s+.*s', '', df$Work.Experience))
df$Work.Experience[ind] = df$Work.Experience[ind] / 12
boxplot(Work.Experience ~ Interview, df, main = 'Interviewed or not vs Work Experience')
boxplot(Work.Experience ~ Hired, df, main = 'Hired or not vs Work.Experience')
```

Based on the boxplots, these hired teachers do not appear to have higher GPA scores.

In addition, these interviewed and hired teachers do not appear to have higher work experience.

In conclusion, there exists an evidence of age discrimination in the interviewing and hiring of teachers.

\newpage 
##Q4

```{r warning=FALSE, echo=FALSE, fig.align='center', fig.height=15, fig.width=12}
df = read.csv('Q4_data.csv', stringsAsFactors = FALSE)
par(mfrow = c(3,2))
tb = table(df$Location)
pie(table(df$Location), col = colours()[3:7], main = 'Pie Chart of Location',
    radius = 1, cex = 1, border = FALSE, cex.main = 2,
    labels = paste0(names(table(df$Location)), ': ', table(df$Location)))
df$Winner = gsub('#[0-9]+\\s+', '', df$Winner)
my_winner = c(table(df$Winner)[1],table(df$Winner)[3],table(df$Winner)[2])
bp <- barplot(my_winner,col = colours()[5:7], ylim = c(0, 80),
        cex.names = 1, ylab = 'Count', main = 'Barplot of Winner', cex.main = 2
      )
text(bp,c(64.5, 56.5, 4.5), my_winner,cex=1.5,pos=3)

df = read.csv('Q4_data.csv', stringsAsFactors = FALSE)
wining_loc_VA<- df[which(df$Winner=='Virginia'),"Location"]
wining_loc_NC<- df[which(df$Winner=='North Carolina'), "Location"]
#par(mfrow=c(1,2))
tb1<-table(wining_loc_VA)
tb2<-table(wining_loc_NC)

pie(tb1,col = colours()[4:7], main="Wining place for VA",
    radius = 1, cex=1, border=FALSE, cex.main=2,
    labels = paste0(names(tb1), ': ', tb1))
?pie
pie(tb2,col = colours()[2:6], main="Wining place for NC",
    radius = 1, cex=1, border=FALSE, cex.main=2,
    labels = paste0(names(tb2), ': ', tb2))


score = as.numeric(unlist(strsplit(df$Score, 'C')))
score1 = score[seq(1, length(score), 2)]
score2 = score[seq(2, length(score), 2)]
score_Vir = ifelse(df$Winner == 'Virginia', score1, score2)
score_Nor = ifelse(df$Winner == 'Virginia', score2, score1)
df['VA_score']<-score_Vir
df['Nor_score']<- score_Nor
df['diff']<- df$VA_score-df$Nor_score
#par(mfrow = c(1,2))
plot(df$diff, type = 'o', col='lightblue', xlab = "Each Competition", 
     ylab = "Score", main = "Score Winning", yaxt="n",
     pch = 19,cex = 0.7, cex.main = 2)
axis(side = 2,at=c(-60,-40,-20,0,20,40,60), labels = c(60,40,20,0,20,40,60))
axis(side = 4, at = c(-30,30), labels = c("NC", "VA"))
abline(h=0, col="pink",lwd=2,lty = 2)

NC_win_score<- df[which(df$diff < 0),]
VA_win_score<- df[which(df$diff > 0),]
data<-c(mean(abs(NC_win_score$diff)), mean(VA_win_score$diff))
bp2<-barplot(data, col=colours()[6:7],
        ylab = "Score wins", names.arg = c("North Carolina", "Virginia"), main = "Average scores winning for all games"
        ,ylim = c(0,25), cex.main = 2)
text(bp,c(18.71875, 15.5), round(data,1), cex = 1.2, pos=3)
```

\newpage
# R Code
##1.
```{r eval=FALSE}
par(mfrow=c(1,1))
digit = matrix(scan('HandWritten.txt'), 16, byrow = TRUE)
image(t(digit[16:1, ]))
```
##2.
```{r eval=FALSE}
df = read.csv('mixture.csv', row.names = 1)
col_list<-ifelse(df$y==1,'blue','red')
plot(x2 ~ x1, df, cex.lab = 1.5, xlim = c(-3.5, 5.5), ylim = c(-3.5, 5.5), bty = 'L',
     col=col_list)
lines(df$x1, (0.978 + 0.134*df$x1)/1.398, lwd = 2)
x1 = sort(df$x1)
x = c(min(x1)-0.5, x1, max(x1)+0.5)
polygon(c(x[1], x, x[length(x)]),
        c(min(df$x2)-0.5, (0.978 + 0.134*x)/1.398, min(df$x2)-0.5), col = "pink")
polygon(c(x[1], x, x[length(x)]),
        c(max(df$x2)+0.5, (0.978 + 0.134*x)/1.398, max(df$x2)+0.5), col = "lightblue")
pchs = ifelse(df$y == 0, 2, 8)
points(df$x1, df$x2, pch = pchs,col=col_list)
legend('topright', legend = c(0, 1), pch = c(2, 8), cex = 1.5, bty = 'n', inset = 0.01
       ,col=c('red','blue'))
```

##3.
```{r eval=FALSE}
df = read.csv('TeacherHires.csv', stringsAsFactors = FALSE, na.strings = 'N/A')
df$Calculated.Age = as.numeric(df$Calculated.Age)
df = df[!is.na(df$Calculated.Age), ]
df$Age = ifelse(df$Calculated.Age < 40, 'young', 'old')
df$Interview = ifelse(df$Interview == 'no', 'no', 'yes')
df$Hired = ifelse(df$Hired == 'no', 'no', 'yes')
par(mfrow = c(1, 2), mar = c(3, 3, 2, 3))
tb1 = round(prop.table(table(df$Interview, df$Age), 2), 4)
barplot(tb1, beside = TRUE, legend.text = TRUE, ylim  = c(0, 1), xlab = 'Age',
        ylab = 'Percentage', main = 'Interviewed or not vs Age',
        args.legend = list(bty = 'n'))
text(c(1.5, 2.5), tb1[, 1]+0.05, paste0(tb1[, 1]*100, '%'))
text(c(4.5, 5.5), tb1[, 2]+0.05, paste0(tb1[, 2]*100, '%'))
tb2 = round(prop.table(table(df$Hired, df$Age), 2), 4)
barplot(tb2, beside = TRUE, legend.text = TRUE, ylim  = c(0, 1), xlab = 'Age',
        ylab = 'Percentage', main = 'Hired or not vs Age',
        args.legend = list(bty = 'n'))
text(c(1.5, 2.5), tb2[, 1]+0.05, paste0(tb2[, 1]*100, '%'))
text(c(4.5, 5.5), tb2[, 2]+0.05, paste0(tb2[, 2]*100, '%'))

df$GPA.undergrad = as.numeric(df$GPA.undergrad)
df$GPA.GRAD = as.numeric(df$GPA.GRAD)
df = df[!((is.na(df$GPA.GRAD)) | (is.na(df$GPA.undergrad))), ]
par(mfrow = c(2, 2), mar = c(3, 3, 3, 2))
boxplot(GPA.undergrad ~ Hired, df, main = 'Hired or not vs GPA Undergrad')
boxplot(GPA.GRAD ~ Hired, df, main = 'Hired or not vs GPA Grad')
ind = grep('months', df$Work.Experience )
df$Work.Experience = as.numeric(gsub('\\s+.*s', '', df$Work.Experience))
df$Work.Experience[ind] = df$Work.Experience[ind] / 12
boxplot(Work.Experience ~ Interview, df, main = 'Interviewed or not vs Work Experience')
boxplot(Work.Experience ~ Hired, df, main = 'Hired or not vs Work.Experience')
```

##4.
```{r eval=FALSE}
df = read.csv('Q4_data.csv', stringsAsFactors = FALSE)
par(mfrow = c(3,2))
tb = table(df$Location)
pie(table(df$Location), col = colours()[3:7], main = 'Pie Chart of Location',
    radius = 1, cex = 1, border = FALSE, cex.main = 2,
    labels = paste0(names(table(df$Location)), ': ', table(df$Location)))
df$Winner = gsub('#[0-9]+\\s+', '', df$Winner)
my_winner = c(table(df$Winner)[1],table(df$Winner)[3],table(df$Winner)[2])
bp <- barplot(my_winner,col = colours()[5:7], ylim = c(0, 80),
        cex.names = 1, ylab = 'Count', main = 'Barplot of Winner', cex.main = 2
      )
text(bp,c(64.5, 56.5, 4.5), my_winner,cex=1.5,pos=3)

df = read.csv('Q4_data.csv', stringsAsFactors = FALSE)
wining_loc_VA<- df[which(df$Winner=='Virginia'),"Location"]
wining_loc_NC<- df[which(df$Winner=='North Carolina'), "Location"]
#par(mfrow=c(1,2))
tb1<-table(wining_loc_VA)
tb2<-table(wining_loc_NC)

pie(tb1,col = colours()[4:7], main="Wining place for VA",
    radius = 1, cex=1, border=FALSE, cex.main=2,
    labels = paste0(names(tb1), ': ', tb1))
?pie
pie(tb2,col = colours()[2:6], main="Wining place for NC",
    radius = 1, cex=1, border=FALSE, cex.main=2,
    labels = paste0(names(tb2), ': ', tb2))


score = as.numeric(unlist(strsplit(df$Score, 'C')))
score1 = score[seq(1, length(score), 2)]
score2 = score[seq(2, length(score), 2)]
score_Vir = ifelse(df$Winner == 'Virginia', score1, score2)
score_Nor = ifelse(df$Winner == 'Virginia', score2, score1)
df['VA_score']<-score_Vir
df['Nor_score']<- score_Nor
df['diff']<- df$VA_score-df$Nor_score
#par(mfrow = c(1,2))
plot(df$diff, type = 'o', col='lightblue', xlab = "Each Competition", 
     ylab = "Score", main = "Score Winning", yaxt="n",
     pch = 19,cex = 0.7, cex.main = 2)
axis(side = 2,at=c(-60,-40,-20,0,20,40,60), labels = c(60,40,20,0,20,40,60))
axis(side = 4, at = c(-30,30), labels = c("NC", "VA"))
abline(h=0, col="pink",lwd=2,lty = 2)

NC_win_score<- df[which(df$diff < 0),]
VA_win_score<- df[which(df$diff > 0),]
data<-c(mean(abs(NC_win_score$diff)), mean(VA_win_score$diff))
bp2<-barplot(data, col=colours()[6:7],
        ylab = "Score wins", names.arg = c("North Carolina", "Virginia"), main = "Average scores winning for all games"
        ,ylim = c(0,25), cex.main = 2)
text(bp,c(18.71875, 15.5), round(data,1), cex = 1.2, pos=3)
```
