---
title: "STAT3280_HW4"
author: "Kaiwen Zhu(kz8pr), Yongyi Li(yl9gq), Xitong Huo(xh2kb)"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1
```{r, echo=FALSE,message=FALSE}
#this part is to print the pictures i created in word document
library(magick)
picture <- image_read('database.png')
print(picture)

```


\newpage
## Q2
Insert all data into our database

```{r, echo=TRUE, results="hide", message=FALSE}
#install all packages we need
library(igraph)
library(stringr)
library(reshape2)
library(dplyr)
library(tidyr)
library(RSQLite)

#Data cleaning
#Read data
cite <- read.table("./statisticians/Data/paperCitAdj.txt", 
                   stringsAsFactors = F)

paper <- scan("./statisticians/Data/paperList.txt",sep="\n",what="")

authors <- read.table("./statisticians/Data/authorList.txt",stringsAsFactors=FALSE)

A2P <- read.table("./statisticians/Data/authorPaperBiadj.txt")

tmp1 <- lapply(paper,function(s)unlist(strsplit(s,'"')))

years_pre <- unlist(lapply(tmp1,function(x) return(trimws(x[5], 
       which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))))[2:3249]

#get the DOI number(primary key)
DOI_pre <- unlist(lapply(tmp1,function(x) return(trimws(x[4], 
       which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))))[2:3249]

#get the title
Title_pre <- unlist(lapply(tmp1,function(x) return(trimws(x[6], 
      which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))))[2:3249]

Cited_count_pre <- unlist(lapply(tmp1,function(x) return(trimws(x[7], 
      which = c("both", "left", "right"), whitespace = "[ \t\r\n]"))))[2:3249]

#get the year
years <- lapply(years_pre, function(x) return (as.numeric(unlist(strsplit(x, ","))[2])))

#get the cited_count
Cited_count <- lapply(Cited_count_pre, function(x) return (as.numeric(unlist(strsplit(x, ","))[2])))

authors <- unlist(lapply(authors,function(x)gsub(" ","_",x)))
A2P <- t(as.matrix(A2P))
colnames(A2P) <- authors
rownames(A2P) <- DOI_pre


my_author_df <- melt(A2P)
#dim(my_author_df)

my_author_df <- subset(filter(my_author_df, value == 1), select = c('Var1', 'Var2'))
publish_id <- seq(1,dim(my_author_df)[1], by=1)

new_df <- my_author_df %>% separate(Var2, c("fname", "lname"), sep = "(_+)(?!.*_)")

#head(my_author_df)
#head(new_df)
#dim(new_df)

#get the list of co_author
co_author <- aggregate(my_author_df[,-1], list(my_author_df[,1]), paste,collapse=', ')

reference_matrix <- t(as.matrix(cite))

colnames(reference_matrix) <- DOI_pre
rownames(reference_matrix) <- DOI_pre

reference_df <- melt(reference_matrix)
reference_df <- subset(filter(reference_df, value == 1), select = c('Var1', 'Var2'))
reference_id <- seq(1, dim(reference_df)[1], by=1)


#get the list of references for each paper
reference <- aggregate(reference_df[,-1], list(reference_df[,1]), paste,collapse=', ')

#Create the R SQL command

db = dbConnect(SQLite(), dbname = "MYSQL.sqlite")
#dbSendQuery(conn = db, "DROP TABLE person")
#dbSendQuery(conn = db, "DROP TABLE reference")

#First table
first_table <- tibble(publish_id,new_df$fname,new_df$lname
                      ,new_df$Var1)
first_table <- setNames(first_table, c("publish_id", "First_name",
                                       "Last_name", "DOI"))

dbSendQuery(db, "CREATE TABLE paper(
                  DOI VARCHAR(50),
                  Year DATE,
                  Title VARCHAR(80),
                  Cited_count SMALLINT,
                  CONSTRAINT pk_paper PRIMARY KEY(DOI))")

dbSendQuery(db, "CREATE TABLE person(
                  publish_id SMALLINT UNSIGNED,
                  First_name VARCHAR(20),
                  Last_name VARCHAR(20),
                  DOI VARCHAR(50),
                  CONSTRAINT pk_person PRIMARY KEY(publish_id),
                  CONSTRAINT fk_person FOREIGN KEY(DOI)
                      REFERENCES paper(DOI))")

dbSendQuery(db, "CREATE TABLE reference(
                  Reference_id SMALLINT UNSIGNED,
                  DOI VARCHAR(50),
                  Refer_DOI VARCHAR(50),
                  CONSTRAINT pk_reference PRIMARY KEY(Reference_id),
                  CONSTRAINT fk_reference FOREIGN KEY(DOI)
                      REFERENCES paper(DOI))")

#Write first table
dbWriteTable(db, "person", first_table,append=TRUE, row.names=FALSE)
#check the name and fields
#dbListTables(db)
#dbListFields(db, "person")

#create the second table
second_table <- tibble(DOI_pre, as.numeric(years),Title_pre, as.numeric(Cited_count))
second_table <- setNames(second_table, c("DOI", "Year", "Title", "Cited_count"))

dbWriteTable(db, "paper", second_table,append=TRUE, row.names=FALSE)
#check names and fields
#dbListTables(db)
dbListFields(db, "paper")

#create the third table
third_table <- tibble(reference_id, reference_df$Var1, reference_df$Var2)
third_table <- setNames(third_table, c("Reference_id", "DOI", "Refer_DOI"))

dbWriteTable(db, "reference", third_table,append=TRUE, row.names=FALSE)
#check names and fields
#dbListTables(db)
#dbListFields(db, "reference")
```

\newpage
## Q3 

Here is th list of author(or co-author) who have ever
published papers with at least four of the following DOI substrings (“10.1214”,“10.1093”,“10.1046”,“10.1111”,“10.1080”,“10.1198”). 

```{r pressure, echo=TRUE}
#Create the database query
query <- dbSendQuery(db, "SELECT First_name, Last_name, COUNT(*)
                          FROM person
                          WHERE (DOI LIKE '%10.1214%')
                          OR (DOI LIKE '%10.1093%')
                          OR (DOI LIKE '%10.1046%')
                          OR (DOI LIKE '%10.1111%')
                          OR (DOI LIKE '%10.1080%')
                          OR (DOI LIKE '%10.1198%')
                          GROUP BY DOI
                          HAVING COUNT(*) >= 4
                     ")

result <- dbFetch(query)
#print the result
result
```


\newpage
## Q4

Here is the research we did:
```{r, echo=TRUE,results="hide", message=FALSE}
#create the database query
query2 <- dbSendQuery(db, "SELECT Refer_DOI, COUNT(*)
                            FROM reference
                            WHERE Refer_DOI IN (
                            SELECT DOI FROM paper WHERE Year < 2010
                            )
                            GROUP BY Refer_DOI
                            ")

result2 <- dbFetch(query2,n=-1)
result2<- setNames(result2, c("DOI", "Citation_Number"))

library(ggplot2)
ggplot(result2) + geom_histogram(aes(x = Citation_Number), na.rm = TRUE)           +scale_x_log10()+scale_y_log10() + ggtitle(label = "The log-frequency vs. The log-citation. ") + xlab("Log-citation") + ylab("Log-frequency")

```
As we can see, after we plot the log-frequency vs. the log-citation, there exists a strong linear relationship between Log-frequency and log-citation. It means that the popularity of an interacting system tend to be scale-free or following a power-law distribution. So, we verify the pattern in the citation system.






